Thesis Title

### 2022-03-20 chapters/paper1.tex
continue results chp 2

Nine body fluid samples and one tissue sample from 9 different patients were sequenced, including one sample from peritoneal fluid, five from pus (3 abscesses and 2 empyemas), two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy (Table \ref{tab:ch2_table_1}).
In total 15 different isolates obtained from the 10 samples were considered of possible clinical significance and were selected for species identification and antimicrobial susceptibility testing during routine work up of the samples (Table \ref{tab:ch2_table2},\ref{} and \ref{}).
In samples 2 and 3, only one colony-forming unit (CFU) of \textit{Escherichia coli} and \textit{Staphylococcus epidermidis}, respectively, was detected after 48 hours of incubation. 
In samples 2 and 5, the anaerobic cultures were mixed to such an extent, that no further characterization of the colonies was performed, and the results were reported as anaerobic mixed culture.

@@

There was fungal growth in 2 samples (1 and 5) that included two Candida species (one \textit{Candida glabrata} and one \textit{Candida albicans}). 
The different bacterial and fungal species identified in each sample are shown in Tables \ref{tab:ch2_table2}, \ref{} and \ref{}.

@@

>>>>

Nine body fluid samples and one tissue sample from 9 different patients were sequenced, including one sample from peritoneal fluid, five from pus (3 abscesses and 2 empyemas), two from synovial fluid of knees with prosthesis, one from sputum and one from a bone biopsy (Table \ref{tab:ch2_table_1}).
In total 15 different isolates obtained from the 10 samples were considered of possible clinical significance and were selected for species identification and antimicrobial susceptibility testing during routine work up of the samples (Table \ref{tab:ch2_table2},\ref{ch2_table3} and \ref{ch2_table4}).
In samples 2 and 3, only one colony-forming unit (CFU) of \textit{Escherichia coli} and \textit{Staphylococcus epidermidis}, respectively, was detected after 48 hours of incubation. 
In samples 2 and 5, the anaerobic cultures were mixed to such an extent, that no further characterization of the colonies was performed, and the results were reported as anaerobic mixed culture.

@@

There was fungal growth in 2 samples (1 and 5) that included two Candida species (one \textit{Candida glabrata} and one \textit{Candida albicans}). 
The different bacterial and fungal species identified in each sample are shown in Tables \ref{tab:ch2_table2}, \ref{ch2_table3} and \ref{ch2_table4}.

@@

\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in webpages (BaseSpace, Taxonomer and CosmosID).}
\label{tab:ch2_table4}
\begin{tabular}{@{}|l|l|l|l|lllll|@{}}
\multirow{2}{*}{\textbf{Sample number}} &
  \multirow{2}{*}{\textbf{Culture result (CFU)$^a$}} &
  \multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}} &
  \multirow{2}{*}{\textbf{WGS-based identification}} &
  \multicolumn{5}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){5-9} 
 &
   &
   &
   &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Genius\\  (Basespace)$^c$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Kraken\\  (Basespace)$^c,d$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}MetaPhlAn\\  (Basespace)$^c$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{Taxonomer (Utah)$^b,e$}} &
  \textbf{Cosmos ID$^a$} \\ \midrule
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (14.4\%)\\  \textit{S. haemolyticus} (55.8\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (25.0\%)\\  \textit{S. haemolyticus} (20.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (65.1\%)\\  \textit{S. haemolyticys} (30.4\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (22.9\%)\\  \textit{S. haemolyticus} (20.1\%)\\  Not identified$^*$\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (50.3\%)\\  \textit{S. haemolyticus} (22.1\%)\\  \textit{C. glabrata} (88.6\%)\end{tabular} \\ \midrule
  \begin{tabular}[c]{@{}l@{}}103\\  1\\  Not determined\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (94.0\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (27.0\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (54.2\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (14.2\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100\%)\end{tabular} \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (0.1\%)} &
  \multicolumn{1}{l|}{\textit{S. pseudintermedius} (3.4\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (0.3\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (8.3\%)} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (0.4\%)\\  Not identified$^*$\\  \textit{S. anginosus} (0.03\%)\\  \textit{E. faecalis} (0.8\%)\\  Several species (45.0\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (10.2\%)\\  \textit{K. oxytoca} (0.5\%)\\  \textit{S. anginosus} (0.4\%)\\  \textit{E. faecalis} (0.3\%)\\  Several species (8.0\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (7.0\%)\\  \textit{K. pneumoniae} (0.01\%)\\  \textit{S. anginosus} (0.3\%)\\  \textit{E. faecalis} (0.7\%)\\  Several species (89.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (3.6\%)\\  \textit{K. michiganensis} (0.1\%)\\  \textit{S. anginosus} (0.1\%)\\  \textit{E. faecalis} (0.1\%)\\  Several species (60.3\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli} (7.6\%)\\  \textit{K. oxytoca} (1.7\%)\\  \textit{S. anginosus} (0.09\%)\\  \textit{E. faecalis} (3.7\%)\\  Several species (86.2\%)\\  Not identified$^*$\end{tabular} \\ \midrule
  \multicolumn{1}{l|}{\textit{E. faecium} (4.2\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (14.8\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (5.5\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (1.4\%)} &
  \textit{E. faecium} (4.1\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (93.8\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (14.2\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (100\%)} &
  \multicolumn{1}{l|}{\textit{O. nthropic} (88.9\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (99.8\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (13.1\%)} &
  \textit{O. intermedium} (49.5\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (99.5\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (12.7\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (32.5\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (94.8\%)} &
  \multicolumn{1}{l|}{\textit{Serratia} spp. (100\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (1.4\%)} &
  S\textit{. marscescens} (38.4\%) \\ \bottomrule
\itel $^e$Full Analysis mode was used 

\subsection{Comparison of standard procedures and shotgun metagenomics for the identification of clinically relevant pathogens}

The tools used for taxonomic classification are shown in Figure \ref{}. 
The total number of reads and the total number of reads mapped against the human genome (hg19) varied between samples, ranging from 3.5\% to 98.9\% (Table \ref{tab:ch2_table_1}). 
The abundance of human reads was not determined by the type of sample but was probably influenced by individual characteristics of each sample and the success of the methods used in depleting the human DNA. 
We identified the microorganisms present using different taxonomical methods, including three Unix-based tools (Kraken, Metaphlan2 and MIDAS), web-based tools including both commercial and freely available solutions (BaseSpace, Taxonomer and CosmosID) and one commercial approach having a graphical interface (CLC Genomics Workbench v10.0.1). 
The taxonomic classification results for each sample are presented in Tables \ref{tab:ch2_table2}, \ref{tab:ch2_table3} and \ref{tab:ch2_table4}. 
In 8 samples, all the microorganisms identified by classical culture were also identified through metagenomics (using at least one method). 
In sample 2, two of the bacterial species identified by classical culture, i.e., \textit{E. coli} and one \textit{Enterococcus avium} were not identified through shotgun metagenomics and in sample 3 there was no concordance between the results of MALDI-TOF and the taxonomical classification methods at the species level (Tables \ref{tab:ch2_table2}, \ref{tab:ch2_table3} and \ref{tab:ch2_table4}).
We identified \textit{Ochrobactrum intermedium} in the negative control, but in low amounts (1.0\% of the reads mapped to the reference genome with the accession number NZ\_ACQA01000002 and only 1.4\% of the reference genome was covered). 
The sensitivity and positive predictive value of each classification method is shown in Table \ref{ch2_table5}.

\begin{table}[]
\caption{Performance of the different taxonomic classification methods for each sample. Sensitivity and positive predictive value were calculated using culture/MALDI-TOF as standards}
\label{tab:ch2_table5}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lllllll@{}}
\toprule
\textbf{Method} &
  \textbf{Total number of bacteria identified$^a$} &
  \textbf{True positives$^a$} &
  \textbf{False positives} &
  \textbf{False negatives} &
  \textbf{Sensitivity (\%)} &
  \textbf{PPV (\%)} \\ \midrule
Culture/MALDI-TOF                                         & 9    & 9 & 0    & 0 & 1    & 1    \\
MetaPhlAn (BaseSpace)                                     & 16   & 7 & 9    & 2 & 0.78 & 0.44 \\
Genius (BaseSpace)                                        & 35   & 8 & 27   & 1 & 0.89 & 0.23 \\
Kraken (BaseSpace)                                        & 959  & 7 & 952  & 2 & 0.78 & 0.01 \\
Taxonomer (Full Analysis)                                 & 4649 & 8 & 4641 & 1 & 0.89 & 0    \\
CosmosID                                                  & 35   & 8 & 27   & 1 & 0.89 & 0.23 \\
Taxonomic Profiling (CLC Genomics Workbench v10.0.1)      & 17   & 6 & 11   & 3 & 0.67 & 0.35 \\
Best match K-mer spectra (CLC Genomics Workbench v10.0.1) & 12   & 8 & 4    & 1 & 0.89 & 0.67 \\
Kraken (Unix)                                             & 198  & 7 & 191  & 2 & 0.78 & 0.04 \\
MetaPhlAn2 (Unix)                                         & 15   & 7 & 6    & 4 & 0.75 & 0.75 \\
MIDAS (Unix)                                              & 34   & 7 & 26   & 2 & 0.88 & 0.5 \\ \bottomrule
\end{tabular}%
}
\tiny
\item $^a$aExcluding the samples with non-identified anaerobic bacteria (Samples 2 and 5)
\item Abbreviations: PPV – positive predictive value
\end{table}
